---
title: "Practical Machine Learning Project"
output:
  html_document:
    fig_height: 7
    fig_width: 10
---

#Introduction
In this project, the goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. The goal of the project is to predict the manner in which they did the exercise. The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. 

#Reading of Data
```{r,message=FALSE,cache=TRUE}
traindata = read.csv("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv")
testdata=read.csv("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv")
```

#Manipulation of Data
Remove index column and columns with NA.
```{r}
traindata=traindata[,c(2:11,37:49,60:68,84:86,160)]
testdata=testdata[,c(2:11,37:49,60:68,84:86)]
```

#Cross Validation
Partition the training data into 60% training and 40% testing for cross validation.
```{r,message=FALSE,warning=FALSE}
set.seed(1753)
library(caret)
in.train <- createDataPartition(traindata$classe, p=0.60, list=FALSE)
train <- traindata[in.train,]
test <- traindata[-in.train,]
```

#Predicting with Tree
```{r,message=FALSE,warning=FALSE}
modFitTree=train(classe~., method="rpart",data=train)
print(modFitTree)
plot(modFitTree$finalModel,main="Classification Tree")
text(modFitTree$finalModel,use.n = T,all=T,cex=0.6)

predictTree <- predict(modFitTree, newdata=test)
confusionMatrix(test$classe,predictTree)
```

#Predicting with Random Forest
```{r,message=FALSE,warning=FALSE}
library(randomForest)
modFitRF=randomForest(classe~.,data=train)
predictRF=predict(modFitRF,newdata=test)
confusionMatrix(test$classe,predictRF)
```

#Conclusion
Looking at the confusion matrix and statistics, it is obvious that the random forest model is a much better prediction model with accuracy of 0.9987



